---
title: "Tutorial on creating maps of Poland"
author: "Maciej Beręsewicz"
date: "`r Sys.Date()`"
output:  rmarkdown::html_vignette
vignette: 
    \VignetteIndexEntry{Tutorial on creating maps of Poland}
    \VignetteEngine{knitr::rmarkdown}
    \VignetteEncoding{UTF-8}
---

Vignettes are long form documentation commonly included in packages. Because they are part of the distribution of the package, they need to be as compact as possible. The `html_vignette` output type provides a custom style sheet (and tweaks some options) to ensure that the resulting html is as small as possible. The `html_vignette` format:

- Never uses retina figures
- Has a smaller default figure size
- Uses a custom CSS stylesheet instead of the default Twitter Bootstrap style

## Vignette Info

Note the various macros within the `vignette` setion of the metadata block above. These are required in order to instruct R how to build the vignette. Note that you should change the `title` field and the `\VignetteIndexEntry` to match the title of your vignette.

## Introduction

### Country level maps

Code below allows to plot maps at country level. In this example we have LAU1 level with NTS2 level borders.

```{r lau1, fig.height=6,fig.width=6,dev='CairoPNG',warnings=FALSE,message=FALSE,out.height=640,out.width=640}
library(MapOfPoland)
par(mar = c(1, 1, 1, 1) + 0.1)
plot(LAU1)
plot(NTS2,lwd=2,add=T)
```

Code below allows to plot maps at country level. In this example we have LAU1 level with NTS2 level borders.

```{r lau2, fig.height=6,fig.width=6,dev='CairoPNG',warnings=FALSE,message=FALSE,out.height=640,out.width=640}
par(mar = c(1, 1, 1, 1) + 0.1)
plot(LAU2)
plot(LAU1,lwd=1,border='red',add=T)
plot(NTS2,lwd=2,add=T)
```


### LAU1 level maps

Now our goal is to choose specific NTS2 level and plot alongside LAU1 and LAU2 level.

```{r nts2select, fig.height=6,fig.width=6,dev='CairoPNG',warnings=FALSE,message=FALSE,out.height=640,out.width=640}
par(mar = c(1, 1, 1, 1) + 0.1)
lau2_level <- substr(LAU2$code,1,2)=='30'
lau1_level <- substr(LAU1$code,1,2)=='30'
nts2_level <- substr(NTS2$code,1,2)=='30'

plot(LAU2[lau2_level,])
plot(LAU1[lau1_level,],lwd=1,border='red',add=T)
plot(NTS2,lwd=2,add=T)
```

### City level maps

Now we would like to plot Poznan at Cadastral 1 and 2 level.

```{r poznan, fig.height=6,fig.width=6,dev='CairoPNG',warnings=FALSE,message=FALSE,out.height=640,out.width=640}
par(mar = c(1, 1, 1, 1) + 0.1)
cu1_level <- substr(Cadastral1$code,1,4)=='3064'
cu2_level <- substr(Cadastral2$code,1,4)=='3064'

plot(Cadastral2[cu2_level,])
plot(Cadastral1[cu1_level,],lwd=2,add=T)
```


Now we would like to plot Warsaw at Cadastral 1 and 2 level.

```{r warsaw, fig.height=6,fig.width=6,dev='CairoPNG',warnings=FALSE,message=FALSE,out.height=640,out.width=640}
par(mar = c(1, 1, 1, 1) + 0.1)
cu1_level <- substr(Cadastral1$code,1,4)=='1465'
cu2_level <- substr(Cadastral2$code,1,4)=='1465'

plot(Cadastral2[cu2_level,])
plot(Cadastral1[cu1_level,],lwd=2,add=T)
```

As we can see, due to decrase in quality to provide maps in R package we have problem that not all borders overlay as supposed. To avoid this problem we can use `maptools::unionSpatialPolygons` (**DAJE BŁĄD**).


## Examples with `ggplot2`


### Plot using `ggplot2`

First we need to use `ggplot2::fortify` function in order to create `data.frame` object

```{r lau1_ggplot2, fig.height=6,fig.width=6,dev='png',warnings=FALSE,message=FALSE,out.height=640,out.width=640}
library(ggplot2)

LAU1_df <- fortify(LAU1)
NTS2_df <- fortify(NTS2)

par(mar = c(1, 1, 1, 1) + 0.1)
ggplot() +
  geom_path(data= LAU1_df, aes(x = long, y = lat, group = id)) +
  geom_path(data= NTS2_df, aes(x = long, y = lat, group = id), size=2) +
  theme_bw() +
  theme(axis.ticks = element_blank(),
        axis.text = element_blank(),
        axis.title = element_blank(),
        panel.grid = element_blank(),
        panel.border = element_blank())
```



## Choropleth maps with ggplot2

Now let plot data from our last elections using `ggplot2` and `dplyr`. First we need to aggreate data on LAU2 level.

```{r elections, fig.height=6,fig.width=6,dev='CairoPNG',warnings=FALSE,message=FALSE}
library(dplyr)
elections_to_plot <- elections2015 %>%
  dplyr::select(round,code,candidate,votes_valid,number_of_votes) %>%
  mutate(code = substr(code,1,4)) %>%
  group_by(round,code,candidate) %>%
  summarise(total = sum(votes_valid),
            votes = sum(number_of_votes),
            frac = votes/total) %>%
  mutate(candidate = factor(
    x = candidate,
    levels = paste0('K',1:11),
    labels = c('Grzegorz Michał Braun',
               'Andrzej Sebastian Duda',
               'Adam Sebastian Jarubas',
               'Bronisław Maria Komorowski',
               'Janusz Ryszard Korwin-Mikke',
               'Marian Janusz Kowalski',
               'Paweł Piotr Kukiz',
               'Magdalena Agnieszka Ogórek',
               'Janusz Marian Palikot',
               'Paweł Jan Tanajno',
               'Jacek Wilk')
  )) %>%
  left_join(
    as.data.frame(LAU1) %>%
      dplyr::select(id,code,name) %>%
      mutate(code = as.character(code)),
    by = c('code'='code')
  )

elections_to_plot_duda_komorowski <- elections_to_plot %>%
  filter(grepl('Duda|Komorowski',candidate))
```

Now plot the data

```{r plot_elections, fig.height=8,fig.width=8,dev='png',warnings=FALSE,message=FALSE,out.height=640,out.width=640}
ggplot() +
  geom_map(data = elections_to_plot_duda_komorowski, 
           aes(map_id = id, fill = frac), map=LAU1_df,colour='black') + 
  geom_path(data = NTS2_df, aes(x = long, y = lat, group = id), size=2) +
  theme_bw() +
  theme(axis.ticks = element_blank(),
        axis.text = element_blank(),
        axis.title = element_blank(),
        panel.grid = element_blank(),
        panel.border = element_blank()) +
  facet_grid(round~candidate) +
  scale_fill_gradient(name='Poparcie',low='white',high='red')
```

Additionally, we can plot this data with scatterplot.

```{r scatter_plot_elections, fig.height=8,fig.width=8,dev='png',warnings=FALSE,message=FALSE,out.height=640,out.width=640}
library(tidyr)
elections_to_plot_duda_komorowski %>%
  dplyr::select(-round,-total,-votes,-id) %>%
  spread(round,frac) %>%
  rename(round1 = `1`,
         round2 = `2`) %>%
  ggplot(data = .,
         aes(x = round1,
             y = round2)) +
  geom_point() +
  facet_wrap(~candidate) +
  geom_abline(a=0,b=1,col='red') + 
  theme_bw() 
```

and finally let see where votes on Kukiz went

```{r scatter_plot_elections2, fig.height=8,fig.width=8,dev='png',warnings=FALSE,message=FALSE,out.height=640,out.width=640}

elections_to_plot %>%
  filter(grepl('Duda|Komorowski|Kukiz',candidate)) %>%
  dplyr::select(-round,-total,-votes,-id) %>%
  ungroup() %>%
  mutate(round = ifelse(grepl('Kukiz',candidate),2,round)) %>%
  spread(candidate,frac) %>%
  filter(round == 2) %>%
  gather(candidate,frac,`Andrzej Sebastian Duda`,`Bronisław Maria Komorowski`) %>%
  ggplot(data = .,
         aes(x = `Paweł Piotr Kukiz`,
             y = frac)) +
  geom_point() +
  facet_wrap(~candidate) +
  geom_abline(a=0,b=1,col='red') + 
  theme_bw() 
```




