---
title: "Tutorial on creating maps of Poland"
author: "Maciej Beręsewicz"
date: "`r Sys.Date()`"
output: 
    rmarkdown::html_vignette
vignette: 
    \VignetteIndexEntry{Tutorial on creating maps of Poland}
    \VignetteEngine{knitr::rmarkdown}
    \VignetteEncoding{UTF-8}
---

Vignettes are long form documentation commonly included in packages. Because they are part of the distribution of the package, they need to be as compact as possible. The `html_vignette` output type provides a custom style sheet (and tweaks some options) to ensure that the resulting html is as small as possible. The `html_vignette` format:

- Never uses retina figures
- Has a smaller default figure size
- Uses a custom CSS stylesheet instead of the default Twitter Bootstrap style

## Vignette Info

Note the various macros within the `vignette` setion of the metadata block above. These are required in order to instruct R how to build the vignette. Note that you should change the `title` field and the `\VignetteIndexEntry` to match the title of your vignette.

## Figures

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r lau1, fig.height=6,fig.width=6,dev='CairoPNG'}
library(MapOfPoland)
par(mar = c(1, 1, 1, 1) + 0.1)
plot(LAU1)
plot(NTS2,lwd=2,add=T)
#par(mar = c(5, 4, 4, 2) + 0.1)
```

You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(elections2015[,1:5], 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))
